A thesis draft

### 2018-11-04 res/sections/08-Background.tex
Add some part on SDN and on virtualization

>>>>

\section{Software-defined networking}
The SDN is an emerging approach to cloud computing that allows to configure
network in a programmatically way, possibly enhancing monitoring and
performances. The main sake of SDN is to the packet forwarding process,
\emph{data plane}, from the routing one, the so called \emph{control plane}.
The latter component serves as the centralized network intelligence and could
consists of one or more controllers. It has some drawbacks in teorms of
security, scalability and elasticity, because of the centralized architecture.

\subsection{The SDN stack}
\begin{figure}[t]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption{SDN architecture}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be splitted into three different layers, based on their
functionalities. \emph{SDN Applications} are programs that communicate the
behaviour to the SDN controllers and request for needed resources and data
through APIs, often called \emph{northbound interfaces}(NBI). NBIs add a level
of abstraction to the higher level application on the set of functionalities of
the controllers: ideally, taking advantage of NBIs, application could use
controllers of different vendors in the same way. So, SDN applications can be
seen as 2 layers programs, the logic and the communication part that uses NBIs.
Theese applications can also build an abstracted view of the network, using
information of controllers, in order to make easier the analysis of the traffic
and simplifing the decision making process.

\emph{SDN controllers}, instead, are logically centralized components that
receive requests from the application and either translate them in a fashion
that can be understood from the lower layer, or collect data needed by the
application. Because of they have to accomplish theese aims, they can be see
as 3 layers programs, in which a layers exposes an interface to the upper
applications (\emph{NBI agents}), the second one represent the SDN controller
logic and finally the third layer has the task to communicate with the network
(Control to Data-Plane Interface or southbound interface).

In the lowest part of the architecture we can find the \emph{SDN Networking Devices}. These ones control packet forwarding and data processing capabilities of the network. They comprises a CDPI agent and a set of one or more traffic forwarding engines and zero or more traffic processing functions (the logic of the devices).

\subsection{The OpenFlow protocol}
The SDN strategy is often associated with OpenFlow protocol: this communication
protocol aim is to give access to the forwarding network component to the lower
layer network infrastructure, made of switches and routers. OpenFlow enables
the remote administration of networks working on layer 3: packets forwarding
tables can be modified relying on packet matching rules. Following this
approach it is possible to change routes periodically or ad-hoc, based on
packet types, traffic or network load.

\begin{figure}[t]
 \centering
 \includegraphics[scale=0.65]{openflow}
 \caption{OpenFlow protocol communication graph}
 \label{chap:background:img:openflow_protocol}
\end{figure}

In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
how switches works in the OpenFlow protocol. Basically, OpenFlow swiches own a
set of tables to organizedpacket control and forwarding. These tables can be
updated by higher-level application, usually an SDN application or controller,
taking advatage of OpenFlow channels (OF channels).

\section{Virtualization}
Cloud computing is now leading-edge Internet technology and virtualization
technology is one of the important part of cloud computing system.
Virtualization, in computing, refers to the act of creating a virtual (rather
than actual) version of something, including but not limited to a virtual
computer hardware platform, operating system (OS), storage device, or computer
network resources. \todo{To change I do not what can I write, stolen from ``The
research and implementation of cloud computing platform based on docker'' paper}

\subsection{Network Virtualization}
%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAccessKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41fBNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_network_virtualization.pdf
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hypervisor, decupling the logical
network from the harwdare one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastucture
(InPs) and who provides the services(SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service. Most known instances of NV are \emph{VPN},
dedicated networks that allows the communication among multiple sites, using
private and secure tunnels, relying on the public network such as Internet.
Other examples of netvork virtualization are \emph{overlay networks}. They are
logical networks, build on the top of an existing physical network, tipically
implemented at application layer.

\subsection{Network Function Virtualization}
The NFC concept refers to the use of virtualization tecnologies to virtualize
network nodes functions. VNFs may run different processes and softwares, on the
top of servers, switches, storage devices or on cloud computing infrastucture,
that substitute custom hardware. That approach worked well on the past, because
of the need to follow rigorous standards, providing at the same time high
quality, but it led to long product cycles and high specialized proprietary
hardware. New Internet senarios, that include IoT and the increasing number of
services to spread multimedia data (Skype, Netflix, Spotify, ...), requires
more flexibles function and faster network deployment. NFV technology is
exploited in a framework, composed at least by three components:
\begin{enumerate}
 \item a software that performs the network function that must be deployed,
 that is the VNF;
 \item NFV Infrastructure (NFVI) that are all hardware and software components
 that build the environment where VNFs are deployed;
 \item NFV Management And Orchestration (NFV-MANO) architectural framework that
 is all functional blocks, data repositories used by these blocks, and
 reference points and interfaces through which these functional blocks exchange
 information for the purpose of managing and orchestrating NFVI and VNFs.
\end{enumerate}
